\documentclass{ximera}

\title{Ratio, root, and comparison tests}

\begin{document}

\begin{abstract}
  For series with positive terms, a series which is smaller term-by-term than a convergent series also converges.
\end{abstract}

\maketitle

We study various convergence tests to determine whether or not a
series converges: in particular, we will consider the comparison test,
the ratio test and the root test.

After working through this material, I hope you will be able to
\begin{itemize}
\item identify a \(p\)-series,
\item compare a series to a \(p\)-series,
\item apply the root test to analyze convergence, and
\item apply the ratio test on a series involving powers,
\item apply the ratio test on a series involving factorials,
\item apply the ratio test on a series involving both factorials and powers,
\end{itemize}

\section{What are \(p\)-series?}

A very important example of a series is a \(p\)-series, which just means a series of the form \[\sum_{n=1}^\infty \frac{1}{n^p}\] for some fixed \(p\).  We can show that a \(p\)-series converges when \(p > 1\) and diverges otherwise.

\youtube{https://www.youtube.com/watch?v=RFB7QSrhgGs}

\begin{question}
  Does the series \(\displaystyle\sum_{m=5}^\infty \left( \displaystyle\frac{7}{4 \, m^{1.64}} \right)\) converge or diverge?
  
  \begin{solution}
    \begin{hint}
      Note that \(\displaystyle\displaystyle\sum_{m=5}^\infty \left( \displaystyle\frac{7}{4 \, m^{1.64}} \right)\) converges iff \(\displaystyle\sum_{m=5} \displaystyle\frac{1}{m^{1.64}}\) converges.
    \end{hint}
    \begin{hint}
      But \(\displaystyle\sum_{m=5} \displaystyle\frac{1}{m^{1.64}}\) is the tail of a \(p\)-series.
    \end{hint}
    \begin{hint}
      In this case, \(p = 1.64\).
    \end{hint}
    \begin{hint}
      Since \(p = 1.64 > 1\), we may conclude that the given series converges.
    \end{hint}


    \begin{multiple-choice}
      \choice[correct]{The series converges.}
      \choice{The series diverges.}
    \end{multiple-choice}

  \end{solution}
\end{question}
            

\section{How can \(p\)-series be used with the comparison test?}

\begin{question}
  Does the series \(\displaystyle\sum_{j=6}^\infty \left( \displaystyle\frac{j - 5}{j^{4.0} + 4 \, j^{3}} \right)\) converge or diverge?

  \begin{solution}
    \begin{hint}
      Set \(a_j = \displaystyle\frac{j - 5}{j^{4.0} + 4 \, j^{3}}\).
    \end{hint}
    \begin{hint}
      Note that \(a_j < \displaystyle\displaystyle\frac{j}{j^{4.0} + 4 \, j^{3}}\).
    \end{hint}
    \begin{hint}
      Furthermore, \(a_j < \displaystyle\displaystyle\frac{j}{j^{4.0}}\) because removing the term from the denominator makes the fraction bigger.
    \end{hint}
    \begin{hint}
      But now \(\displaystyle\displaystyle\frac{j}{j^{4.0}} = \displaystyle\displaystyle\frac{1}{j^{3}}\).
    \end{hint}
    \begin{hint}
      So \(a_j < \displaystyle\displaystyle\frac{1}{j^{3}}\).
    \end{hint}
    \begin{hint}
      But \(\displaystyle\sum_{j = 6}^\infty \displaystyle\displaystyle\frac{1}{j^{3}}\) converges.
    \end{hint}
    \begin{hint}
      And so the original series converges.
    \end{hint}
    

    \begin{multiple-choice}
      \choice[correct]{The series converges.}
      \choice{The series diverges.}
      
    \end{multiple-choice}

  \end{solution}
\end{question}

\hrule

\begin{question}
  Does the series \(\displaystyle\sum_{n=3}^\infty \left( \displaystyle\frac{n + 5}{n^{2} - n} \right)\) converge or diverge?
  
  \begin{solution}
    \begin{hint}
      Set \(a_n = \displaystyle\frac{n + 5}{n^{2} - n}\).
    \end{hint}
    \begin{hint}
      Note that \(a_n > \displaystyle\displaystyle\frac{n}{n^{2} - n}\).
    \end{hint}
    \begin{hint}
      Furthermore, \(a_n > \displaystyle\displaystyle\frac{n}{n^{2}}\) because removing the term from the denominator makes the fraction smaller.
    \end{hint}
    \begin{hint}
      But now \(\displaystyle\displaystyle\frac{n}{n^{2}} = \displaystyle\displaystyle\frac{1}{n}\).
    \end{hint}
    \begin{hint}
      So \(a_n > \displaystyle\displaystyle\frac{1}{n}\).
    \end{hint}
    \begin{hint}
      But \(\displaystyle\sum_{n = 3}^\infty \displaystyle\displaystyle\frac{1}{n}\) diverges.
    \end{hint}
    \begin{hint}
      And so the original series diverges.
    \end{hint}

    \begin{multiple-choice}
      \choice[correct]{The series diverges.}
      \choice{The series converges.}
      
    \end{multiple-choice}
    
  \end{solution}
\end{question}

\end{document}
