\documentclass{ximera}

\title{Root test}

\begin{document}

\begin{abstract}
  The root test begins by considering the limit of the n-th root of the n-th term.
\end{abstract}

\maketitle

\section{What is the root test?}

The root test is somewhat similar to the ratio test.  Again, suppose \(a_n > 0\) and you want to determine whether the series \(\sum_{n=1}^\infty a_n\) converges or diverges.  Compute \(L = \displaystyle\lim_{n\to\infty} \sqrt[n]{a_{n}}\), and if \(L < 1\), the series converges, while if \(L > 1\), the series diverges.  If \(L = 1\), then the test is inconclusive.

It is no secret that I do not think too highly of the root test.

\youtube{https://www.youtube.com/watch?v=a5TGqoge7hA}

Nevertheless, there are situations where the root test proves the convergence of a series for which the ratio test is silent.

% Relevant video: root-test-useless
\begin{question}
  Does the series \(\displaystyle\sum_{n=7}^\infty \left( \left(\displaystyle\frac{5 \, n + 4}{7 \, n + 8}\right)^{n} \right)\) converge or diverge?

  \begin{solution}
    \begin{hint}
      Let's set \(a_n = \left(\displaystyle\frac{5 \, n + 4}{7 \, n + 8}\right)^{n}\).
    \end{hint}
    \begin{hint}
      This series involves powers of \(n\), so the ratio test might be helpful.
    \end{hint}
    \begin{hint}
      So \(\displaystyle\displaystyle\frac{a_{n+1}}{a_n} = \displaystyle\frac{\left(\displaystyle\frac{5 \, n + 9}{7 \, n + 15}\right)^{n + 1}}{\left(\displaystyle\frac{5 \, n + 4}{7 \, n + 8}\right)^{n}}\).
    \end{hint}
    \begin{hint}
      That does not look like a particularly fruitful path to go down; perhaps we could try the root test?
    \end{hint}
    \begin{hint}
      Note that \(\sqrt[n]{a_{n}} = \displaystyle\frac{5 \, n + 4}{7 \, n + 8}\).
    \end{hint}
    \begin{hint}
      We calculate \(\lim_{n \to \infty} \displaystyle\frac{5 \, n + 4}{7 \, n + 8} = \displaystyle\frac{5}{7}\).
    \end{hint}
    \begin{hint}
      So \(L = \displaystyle\frac{5}{7}\).
    \end{hint}
    \begin{hint}
      Since \(L < 1\), we may conclude by the root test that the given series coverges.
    \end{hint}


    \begin{multiple-choice}
      \choice[correct]{The series converges.}
      \choice{The series diverges.}
    \end{multiple-choice}
  \end{solution}

\end{question}
            

\end{document}
